A new tidy data structure to support exploration and modeling of temporal data

### 2019-07-31 reviews/response.Rmd

---
title: Response to reviewers
author: JCGS-19-044
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the reviewers for their helpful feedback which has led to many improvements in the paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

# Reviewer 1 comments {-}

# Reviewer 2 comments {-}

> P5L37 "Temporal data tools need to catch up." This is a little casual. Could perhaps be rephrased as "This paper and the associated **tsibble** R package extends the tidy way of thinking to temporal data."

Done.

> P7L44 "A variable number of time representations can be spotted in the wild." This is a little casual, and I’m not fond of "variable number". Perhaps "Many different representations of temporal data can be spotted in the wild." (The "spotted in the wild" is still a little informal – but I like it.)

Done.

> P7L50 "This diversity and time zone..." suggest rephrase "The tsibble's index column leverages the date-time data type in R to accommodate diverse temporal representations and associated time zone information."

TODO.

> P8L16 "The identifying variables from which the "key" is constituted remain the same as in the original table with no further tweaks." It's not totally clear to me what is meant here. Is it saying that like with grouping columns (i.e. columns supplied to group_by()), key columns cannot be altered in subsequent analyses? It might be useful to make the comparison to group_by() explicit as it also allows multiple fields. The distinction between group_by() and key() is clear in that the collection of columns that constitute a key must uniquely identify observations (obviously this isn’t the case for grouping variables).

TODO.

> P8L55 "fit models that expect **regular** time series." Should regular be added to the sentence.

Done.

> P10L19 "Duplicates signal a data quality..."

Done.

> P12-13 Consider referencing "fable: and possibly "mable" too.

TODO.

> P14-15 I had to read section 5.2 a few times before I understood it... I think what threw me was the last paragraph starts with a reference to purrr, but the functions described in the remainder of the paragraph describes functions in the tsibble package. Perhaps move the purrr sentence up to the previous paragraph when functional programming is first mentioned.

TODO.

> P15L46 "...table verb achieves in collapsing a tsibble..." perhaps "...table verb collapse a tsibble..."

Done.

> P15L53 "On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone" perhaps "The extensible design of the tsibble package means that the table verbs can incorporate many third-party vector functions."

Done.

> Fig 5 y-axis doesn't need the overall label "Delayed" or if it stays, should be renamed “Status”. Similarly for the legend label.

TODO.

> Fig 6 Consider replacing gather() with pivot_longer() which is intended to be easier to use particularly for more complex "gathering" operations. https://tidyr.tidyverse.org/dev/articles/pivot.html

TODO.

> Fig 7 y-axis label should be "Proportion of flights delayed" or similar. Legend label "Originating airport" (but legend can also be removed, because it is covered by the facet labels.

TODO.

> Fig 9 y-label "Departure delay (mins)". X-axis could be converted to time... e.g. ggplot2::scale_x_time() Could introduce some alpha transparency to the geom_line(). Caption text currently states "Passengers are apt to hold up around 18 during a day, and are recommended to travel early. The variations increase substantially as the upper tails." Both sentences need to be reworded.

TODO.

> P22L3 "might be joined" should it be "can be joined" or perhaps a bit more detail, "Aggregating the half-hourly electricity data to the same daily time interval as the temperature data allows us to join the two data tables to explore..."

TODO.

> P22L27 Suggest a reference to Wang, E, D Cook & RJ Hyndman (2018) to explicitly acknowledge that the frame_calendar() function is part of a different package (sugrrants).

TODO.

> P22L37 "degrees C" to "degrees Celsius" or °C

Done.

> Fig 12 remove legend label "hot"

TODO.

> P25L19 It's not "A missing piece of the tsibble package" because later in the paragraph it says that custom calendars are easily embedded in the tsibble framework. I suggest rephrasing the first sentence in a more positive light... "The tsibble package can easily be extended to accommodate user-defined calendars." Suggest also referencing at least one package that provides a user-defined calendar (perhaps a market-specific business calendar).

TODO.

> P26L54 there’s an errant

Fixed.

> P28L22 Journal of Statistical Software

Fixed.

### 2019-07-31 tsibble.Rmd
added reviewer 2 comments

The **tidyverse** largely formalizes exploratory data analysis. Many in the R community have adopted the **tidyverse** way of thinking and extended it to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018]. Temporal data tools need to catch up.

@@

A variable number of time representations can be spotted in the wild. A date-time object, universally accepted across computing systems, is the most commonly used type for representing time. Date-time also typically associates with a time zone including adjustments such as summer time. This diversity and time zone is acknowledged and accommodated by tsibble's index. When creating a tsibble, time indices are arranged from past to future within each series for the strict temporal ordering that is assumed by time series operations.

@@

One of the cornerstones of time series data, and hence beneath a tsibble, is the time interval. This information plays a critical role in computing statistics (e.g. seasonal unit root tests) and building models (e.g. seasonal ARIMA). The principal divide is between regularly or irregularly spaced observations in time. A tsibble permits implicit missing time, making it difficult to distinguish regularity from the index. It relies on a user's specification by switching the `regular` argument off, when the data involves irregular intervals. This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect time series.

@@

Figure \@ref(fig:tsibble-pipeline) illustrates the distinction of a time series pipeline from a regular data pipeline. It is highly recommended to check for identical entries of key and index before constructing a tsibble. Duplicates signal the data quality issue, which would likely affect subsequent analyses and hence decision making. Analysts are encouraged to gaze at data early and reason about the process of data cleaning. When the data meets the tsibble standard, it flows neatly into the analysis stage and takes full advantage of the tsibble infrastructure.

@@

All user-facing functions can be roughly organized into three major chunks according to their functionality: vector functions (1d), table verbs (2d), and window family. Each chunk is an independent module, but works interdependently. Vector functions in the package mostly deal with time. The atomic functions (such as `yearmonth()` and `yearquarter()`) embedded in the `index_by()` table verb achieves in collapsing a tsibble to a less granular interval. The substitution of another time function in the `index_by()` results in the aggregation of different time resolution. Since these time functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages. On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone. It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions, and increase the code readability.

@@

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)). Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees C. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening. This plot illustrates how the tsibble data can easily integrate with other tools and graphics.

>>>>

The **tidyverse** largely formalizes exploratory data analysis. Many in the R community have adopted the **tidyverse** way of thinking and extended it to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018]. This paper and the associated **tsibble** R package extends the tidy way of thinking to temporal data.

@@

Many different representations of time can be spotted in the wild. A date-time object, universally accepted across computing systems, is the most commonly used type for representing time. Date-time also typically associates with a time zone including adjustments such as summer time. This diversity and time zone is acknowledged and accommodated by tsibble's index. When creating a tsibble, time indices are arranged from past to future within each series for the strict temporal ordering that is assumed by time series operations.

@@

One of the cornerstones of time series data, and hence beneath a tsibble, is the time interval. This information plays a critical role in computing statistics (e.g. seasonal unit root tests) and building models (e.g. seasonal ARIMA). The principal divide is between regularly or irregularly spaced observations in time. A tsibble permits implicit missing time, making it difficult to distinguish regularity from the index. It relies on a user's specification by switching the `regular` argument off, when the data involves irregular intervals. This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect regular time series.

@@

Figure \@ref(fig:tsibble-pipeline) illustrates the distinction of a time series pipeline from a regular data pipeline. It is highly recommended to check for identical entries of key and index before constructing a tsibble. Duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making. Analysts are encouraged to gaze at data early and reason about the process of data cleaning. When the data meets the tsibble standard, it flows neatly into the analysis stage and takes full advantage of the tsibble infrastructure.

@@

All user-facing functions can be roughly organized into three major chunks according to their functionality: vector functions (1d), table verbs (2d), and window family. Each chunk is an independent module, but works interdependently. Vector functions in the package mostly deal with time. The atomic functions (such as `yearmonth()` and `yearquarter()`) embedded in the `index_by()` table verb collapses a tsibble to a less granular interval. The substitution of another time function in the `index_by()` results in the aggregation of different time resolution. Since these time functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages. The extensible design of the tsibble package means that the table verbs can incorporate many third-party vector functions. It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions, and increase the code readability.

@@

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)). Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees Celsius. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening. This plot illustrates how the tsibble data can easily integrate with other tools and graphics.
